A new software tool to facilitate environmental modelling procedures based on meteorological data management

### 2020-04-21 latex_manuscript/elsevier/6_Sustainable_Energy_Technologies_and_Assessments/2 - envio/spamda.tex
Envío a la revista: Sustainable_Energy_Technologies_and_Assessments

		\begin{abstract}
			Meteorological data are extensively used to perform environmental learning. Machine Learning (ML) techniques, a valuable support in many research areas, require datasets containing information related to the topic under study, which are not always available in an appropriate format, and its preparation and pre-processing implies a lot of time and effort by the researchers. This paper presents a novel software tool with an user-friendly GUI to create datasets by means of management and data integration of meteorological observations from two well-known sources of information: the \textit{National Data Buoy Center} and the \textit{National Centers for Environmental Prediction and for Atmospheric Research Reanalysis Project}. Such datasets can be created using buoys and reanalysis data by customisables procedures, in terms of predictive and objective variables and temporal resolution, to be used with ML methodologies for prediction tasks (classification or regression) that support improvement of sustainability energy production, design of production systems as WECs or environmental modelling, among others.
		\end{abstract}

@@

		Up to now, and to the best of our knowledge, this is the first software tool addressing the problem previously discussed, integrating meteorological data from NDBC and NNRP. The novelties and functionalities that SPAMDA offers to the researchers will be detailed in Section \ref{sec:SPAMDA}, although some of them are briefly summarised below:
		\begin{itemize}

			\item The generation of datasets becomes a very easy and customisable task, by means of the selection of different input parameters, such as predictive and objective variables, classification and regression ML techniques, output discretisation or prediction horizon, among others.
			\item It makes the researcher focus on environmental modelling, without having to worry about mechanical tasks, avoiding laborious and pre-processing procedures that implies a lot of time and effort.
			\item It manages the extensive casuistry of the data integration which can lead to incomplete datasets.
			\item It avoids possible researcher errors in the intermediate steps of the process of creation of the datasets, such as time and geographical coordinates conversion or missing values handling (dates or measurements not recorded), among others.
			\item It provides information about the quality and quantity of the data.
			\item It includes different pre-processing tasks, such as normalisation and missing data recovery.
			\item It facilitates data management and well-organised storage of the datasets.
			\item The created datasets can be easily used by ML tools.
			\item Its modular design allows the implementation of new functional modules for managing meteorological data from others sources for renewable energy and environmental research.
			\item It includes an user-friendly GUI, facilitating and greatly simplifying data management, and it is integrated with the Explorer environment of WEKA.
			\item It is multi-platform, and it can be used on any computer with Java regardless of the operating system.
		\end{itemize}

@@

		A new open source tool named SPAMDA with an user-friendly GUI for creating datasets using meteorological data from NDBC and NNRP has been presented in this work. These datasets, that will be ready to use as input for ML techniques in prediction tasks (classification or regression), are easily obtained by means of the selection of different input parameters, such as predictive and objective variables, output discretisation or prediction horizon. As a result, the researchers will benefit from significant support when carrying out environmental modelling related to energy, atmospheric or oceanic studies, among others. Moreover, given that SPAMDA simplifies all the intermediate steps involved in the creation of datasets and manages the extensive casuistry of the data integration (such as entering the meteorological information, managing with the incomplete data, pre-processing tasks, the customisable matching process to merge the data and the preparation of the datasets according to the ML technique to use), it avoids errors and reduces the time needed. In this way, the researchers will be able to have more in-depth analysis, which could result in more complete conclusions about the issue under study.

>>>>

		\begin{abstract}
			Meteorological data are extensively used to perform environmental learning. Machine Learning (ML) techniques, a valuable support in many research areas, require datasets containing information related to the topic under study, which are not always available in an appropriate format, and its preparation and pre-processing implies a lot of time and effort by the researchers. This paper presents a novel software tool with an user-friendly GUI to create datasets by means of management and data integration of meteorological observations from two well-known sources of information: the \textit{National Data Buoy Center} and the \textit{National Centers for Environmental Prediction and for Atmospheric Research Reanalysis Project}. Such datasets can be created using buoys and reanalysis data by customisable procedures, in terms of predictive and objective variables and temporal resolution. These datasets can be used by ML methodologies for prediction tasks (classification or regression) that support improvement of sustainability energy production, design of production systems as WECs or environmental modelling, among others.
		\end{abstract}

@@

		Up to now, and to the best of our knowledge, this is the first software tool addressing the problem previously discussed, integrating meteorological data from NDBC and NNRP. The novelties and functionalities that SPAMDA offers to the researchers will be detailed in Section \ref{sec:SPAMDA}, although some of them are briefly summarised below:
		\begin{itemize}
			\item The generation of datasets becomes a very easy and customisable task, by means of the selection of different input parameters, such as predictive and objective variables, classification and regression ML techniques, output discretisation or prediction horizon, among others.
			\item It makes the researcher focus on environmental modelling, without having to worry about mechanical tasks, avoiding laborious pre-processing procedures that imply a lot of time and effort.
			\item It manages the extensive casuistry of data integration which can lead to incomplete datasets.
			\item It avoids possible researcher errors in the intermediate steps of the process, such as time and geographical coordinates conversion or missing values handling (dates or measurements not recorded), among others.
			\item It provides information about the quality and quantity of the data.
			\item It includes different pre-processing tasks, such as normalisation and missing data recovery.
			\item It facilitates data management and well-organised storage of the datasets.
			\item The created datasets can be easily used by ML tools.
			\item Its modular design allows the implementation of new modules for managing meteorological data from others sources, benefiting future renewable energy and environmental research.
			\item It includes a user-friendly GUI, facilitating and greatly simplifying data management, and it is integrated with the Explorer environment of WEKA.
			\item It is multi-platform, and it can be used on any computer with Java regardless of the operating system.
		\end{itemize}

@@

		A new open source tool named SPAMDA with an user-friendly GUI for creating datasets using meteorological data from NDBC and NNRP has been presented in this work. These datasets, that will be ready to be used as input for ML techniques in prediction tasks (classification or regression), are easily obtained by means of the selection of different input parameters, such as predictive and objective variables, output discretisation or prediction horizon. As a result, the researchers will benefit from significant support when carrying out environmental modelling related to energy, atmospheric or oceanic studies, among others. Moreover, given that SPAMDA simplifies all the intermediate steps involved in the creation of datasets and manages the extensive casuistry of the data integration (such as entering the meteorological information, managing with the incomplete data, pre-processing tasks, the customisable matching process to merge the data and the preparation of the datasets according to the ML technique to use), it avoids errors and reduces the time needed. In this way, the researchers will be able to have more in-depth analysis, which could result in more complete conclusions about the issue under study.
